---
title: "R Notebook"
output: html_notebook
---
Run `download_data.Rmd` and `percentage_of_regional_richness.Rmd` First!

```{r}
city_data
```


```{r}
fetch_city_data_for <- function(pool_name, include_city_name = F, include_pool_size = T) {
  results_filename <- paste(paste(pool_name, 'city', 'richness', 'intercept', sep = "_"), "csv", sep = ".")
  results <- read_csv(results_filename)
  
  joined <- left_join(city_data, results)
  
  pool_size_col_name <- paste(pool_name, 'pool', 'size', sep = "_")
  
  required_columns <- c("population_growth", "rainfall_monthly_min", "rainfall_annual_average", "rainfall_monthly_max", "temperature_annual_average", "temperature_monthly_min", "temperature_monthly_max", "happiness_negative_effect", "happiness_positive_effect", "happiness_future_life", "number_of_biomes", "realm", "biome_name", "region_20km_includes_estuary", "region_50km_includes_estuary", "region_100km_includes_estuary", "city_includes_estuary", "region_20km_average_pop_density", "region_50km_average_pop_density", "region_100km_average_pop_density", "city_max_pop_density", "city_average_pop_density", "mean_population_exposure_topm2_5_2019", "region_20km_cultivated", "region_20km_urban", "region_50km_cultivated", "region_50km_urban", "region_100km_cultivated", "region_100km_urban", "region_20km_elevation_delta", "region_20km_mean_elevation", "region_50km_elevation_delta", "region_50km_mean_elevation", "region_100km_elevation_delta", "region_100km_mean_elevation", "city_elevation_delta", "city_mean_elevation", "urbanPshrubs", "permanent_water", "open_forest", "herbaceous_wetland", "herbaceous_vegetation", "cultivated", "closed_forest", "share_of_population_within_400m_of_open_space", "percentage_urban_area_as_streets", "percentage_urban_area_as_open_public_spaces_and_streets", "percentage_urban_area_as_open_public_spaces", "city_gdp_per_population", "city_ndvi", "city_ssm", "city_susm", "region_20km_ndvi", "region_20km_ssm", "region_20km_susm", "region_50km_ndvi", "region_50km_ssm", "region_50km_susm", "region_100km_ndvi", "region_100km_ssm", "region_100km_susm", "city_percentage_protected", "region_20km_percentage_protected", "region_50km_percentage_protected", "region_100km_percentage_protected")
  
  if (include_pool_size) {
    required_columns <- append(c(pool_size_col_name), required_columns)
  }
  
  if (include_city_name) {
    required_columns <- append(c("name"), required_columns)
  }
  
  required_columns <- append(c("response"), required_columns)
  
  joined[,required_columns]
}
```


```{r}
merlin_city_data <- fetch_city_data_for('merlin')
merlin_city_data
```

```{r}
library(randomForest)
library(reshape2)
library(rpart)
library(ggplot2)
library(tidyverse)

library(multcomp)
library(car)
```

```{r}
merlin_city_data_fixed <- rfImpute(response ~ ., merlin_city_data)
merlin_city_data_fixed
```

```{r}
ggplot(merlin_city_data_fixed, aes(response)) + geom_histogram(binwidth = 2)
```

```{r}
source('./random_forest_selection_functions.R')
```

```{r}
scale_parameter_name <- function(scale, postscript) {
  paste('region', paste(scale, 'km', sep = ''), postscript, sep = '_')  
}

scale_parameters <- function(postscript) {
  c(scale_parameter_name(20, postscript), scale_parameter_name(50, postscript), scale_parameter_name(100, postscript))
}

scales_parameters_without <- function(scale_to_exclude, postscript) {
  scales <- scale_parameters(postscript)
  scales[scales != scale_parameter_name(scale_to_exclude, postscript)]
}

select_scales <- function(urban, cultivated, elevation_delta, mean_elevation, average_pop_density, includes_estuary, ssm, susm, ndvi, percentage_protected) {
  append(
    append(
      append(
        append(
          scales_parameters_without(scale_to_exclude = urban, postscript = 'urban'),
          scales_parameters_without(scale_to_exclude = cultivated, postscript = 'cultivated')
        ),
        append(
          scales_parameters_without(scale_to_exclude = elevation_delta, postscript = 'elevation_delta'),
          scales_parameters_without(scale_to_exclude = mean_elevation, postscript = 'mean_elevation')
        )
      ),
      append(
        append(
          scales_parameters_without(scale_to_exclude = average_pop_density, postscript = 'average_pop_density'),
          scales_parameters_without(scale_to_exclude = includes_estuary, postscript = 'includes_estuary')
        ),
        append(
          scales_parameters_without(scale_to_exclude = ssm, postscript = 'ssm'),
          scales_parameters_without(scale_to_exclude = susm, postscript = 'susm')
        )
      )
    ),
    append(
      scales_parameters_without(scale_to_exclude = ndvi, postscript = 'ndvi'),
      scales_parameters_without(scale_to_exclude = percentage_protected, postscript = 'percentage_protected')
    )
  )
}
```

```{r}
select_scales(urban = 20, cultivated = 100, elevation_delta = 20, mean_elevation = 100, average_pop_density = NA, includes_estuary = NA, ssm = 20, susm = 20, ndvi = 100, percentage_protected = NA)
```

select_scales(urban = , cultivated = , elevation_delta = , mean_elevation = , average_pop_density = , includes_estuary = , ssm = , susm = , ndvi =, percentage_protected = )

```{r}
select_variables_from_random_forest(merlin_city_data_fixed)
```

```{r}
exclude_merlin <- !names(merlin_city_data_fixed) %in% select_scales(urban = 20, cultivated = 100, elevation_delta = 50, mean_elevation = 20, average_pop_density = 50, includes_estuary = NA, ssm = 100, susm = 50, ndvi = 20, percentage_protected = 50)

merlin_city_data_fixed_single_scale <- merlin_city_data_fixed[,exclude_merlin]
merlin_city_data_fixed_single_scale
```

```{r}
select_variables_from_random_forest(merlin_city_data_fixed_single_scale)
```

```{r}
create_fifty_rows_of_oob(merlin_city_data_fixed[,c("response", "merlin_pool_size")])
create_fifty_rows_of_oob(merlin_city_data_fixed[,c("response", "merlin_pool_size", "biome_name")])
create_fifty_rows_of_oob(merlin_city_data_fixed[,c("response", "merlin_pool_size", "biome_name", "realm")])
create_fifty_rows_of_oob(merlin_city_data_fixed[,c("response", "merlin_pool_size", "biome_name", "realm", "region_100km_ssm")])
create_fifty_rows_of_oob(merlin_city_data_fixed[,c("response", "merlin_pool_size", "biome_name", "realm", "region_100km_ssm", "temperature_annual_average")])
create_fifty_rows_of_oob(merlin_city_data_fixed[,c("response", "merlin_pool_size", "biome_name", "realm", "region_100km_ssm", "temperature_annual_average", "temperature_monthly_min")])
create_fifty_rows_of_oob(merlin_city_data_fixed[,c("response", "merlin_pool_size", "biome_name", "realm", "region_100km_ssm", "temperature_annual_average", "temperature_monthly_min", "region_50km_elevation_delta")])
create_fifty_rows_of_oob(merlin_city_data_fixed[,c("response", "merlin_pool_size", "biome_name", "realm", "region_100km_ssm", "temperature_annual_average", "temperature_monthly_min", "region_50km_elevation_delta",  "rainfall_monthly_min")])
create_fifty_rows_of_oob(merlin_city_data_fixed[,c("response", "merlin_pool_size", "biome_name", "realm", "region_100km_ssm", "temperature_annual_average", "temperature_monthly_min", "region_50km_elevation_delta",  "rainfall_monthly_min", "permanent_water")])
create_fifty_rows_of_oob(merlin_city_data_fixed[,c("response", "merlin_pool_size", "biome_name", "realm", "region_100km_ssm", "temperature_annual_average", "temperature_monthly_min", "region_50km_elevation_delta",  "rainfall_monthly_min", "permanent_water", "region_50km_susm")])
create_fifty_rows_of_oob(merlin_city_data_fixed[,c("response", "merlin_pool_size", "biome_name", "realm", "region_100km_ssm", "temperature_annual_average", "temperature_monthly_min", "region_50km_elevation_delta",  "rainfall_monthly_min", "permanent_water", "region_50km_susm", "region_20km_ndvi")])
create_fifty_rows_of_oob(merlin_city_data_fixed[,c("response", "merlin_pool_size", "biome_name", "realm", "region_100km_ssm", "temperature_annual_average", "temperature_monthly_min", "region_50km_elevation_delta",  "rainfall_monthly_min", "permanent_water", "region_50km_susm", "region_20km_ndvi", "region_20km_urban")])
create_fifty_rows_of_oob(merlin_city_data_fixed[,c("response", "merlin_pool_size", "biome_name", "realm", "region_100km_ssm", "temperature_annual_average", "temperature_monthly_min", "region_50km_elevation_delta",  "rainfall_monthly_min", "permanent_water", "region_50km_susm", "region_20km_ndvi", "region_20km_urban", "shrubs")])
create_fifty_rows_of_oob(merlin_city_data_fixed[,c("response", "merlin_pool_size", "biome_name", "realm", "region_100km_ssm", "temperature_annual_average", "temperature_monthly_min", "region_50km_elevation_delta",  "rainfall_monthly_min", "permanent_water", "region_50km_susm", "region_20km_ndvi", "region_20km_urban", "shrubs",  "city_gdp_per_population")])
create_fifty_rows_of_oob(merlin_city_data_fixed[,c("response", "merlin_pool_size", "biome_name", "realm", "region_100km_ssm", "temperature_annual_average", "temperature_monthly_min", "region_50km_elevation_delta",  "rainfall_monthly_min", "permanent_water", "region_50km_susm", "region_20km_ndvi", "region_20km_urban", "shrubs",  "city_gdp_per_population", "happiness_positive_effect")])
create_fifty_rows_of_oob(merlin_city_data_fixed[,c("response", "merlin_pool_size", "biome_name", "realm", "region_100km_ssm", "temperature_annual_average", "temperature_monthly_min", "region_50km_elevation_delta",  "rainfall_monthly_min", "permanent_water", "region_50km_susm", "region_20km_ndvi", "region_20km_urban", "shrubs",  "city_gdp_per_population", "happiness_positive_effect", "city_percentage_protected")])
create_fifty_rows_of_oob(merlin_city_data_fixed[,c("response", "merlin_pool_size", "biome_name", "realm", "region_100km_ssm", "temperature_annual_average", "temperature_monthly_min", "region_50km_elevation_delta",  "rainfall_monthly_min", "permanent_water", "region_50km_susm", "region_20km_ndvi", "region_20km_urban", "shrubs",  "city_gdp_per_population", "happiness_positive_effect", "city_percentage_protected", "region_100km_cultivated")])
create_fifty_rows_of_oob(merlin_city_data_fixed[,c("response", "merlin_pool_size", "biome_name", "realm", "region_100km_ssm", "temperature_annual_average", "temperature_monthly_min", "region_50km_elevation_delta",  "rainfall_monthly_min", "permanent_water", "region_50km_susm", "region_20km_ndvi", "region_20km_urban", "shrubs",  "city_gdp_per_population", "happiness_positive_effect", "city_percentage_protected", "region_100km_cultivated", "share_of_population_within_400m_of_open_space")])
create_fifty_rows_of_oob(merlin_city_data_fixed[,c("response", "merlin_pool_size", "biome_name", "realm", "region_100km_ssm", "temperature_annual_average", "temperature_monthly_min", "region_50km_elevation_delta",  "rainfall_monthly_min", "permanent_water", "region_50km_susm", "region_20km_ndvi", "region_20km_urban", "shrubs",  "city_gdp_per_population", "happiness_positive_effect", "city_percentage_protected", "region_100km_cultivated", "share_of_population_within_400m_of_open_space", "rainfall_monthly_max")])
create_fifty_rows_of_oob(merlin_city_data_fixed[,c("response", "merlin_pool_size", "biome_name", "realm", "region_100km_ssm", "temperature_annual_average", "temperature_monthly_min", "region_50km_elevation_delta",  "rainfall_monthly_min", "permanent_water", "region_50km_susm", "region_20km_ndvi", "region_20km_urban", "shrubs",  "city_gdp_per_population", "happiness_positive_effect", "city_percentage_protected", "region_100km_cultivated", "share_of_population_within_400m_of_open_space", "rainfall_monthly_max", "city_ndvi")])
create_fifty_rows_of_oob(merlin_city_data_fixed[,c("response", "merlin_pool_size", "biome_name", "realm", "region_100km_ssm", "temperature_annual_average", "temperature_monthly_min", "region_50km_elevation_delta",  "rainfall_monthly_min", "permanent_water", "region_50km_susm", "region_20km_ndvi", "region_20km_urban", "shrubs",  "city_gdp_per_population", "happiness_positive_effect", "city_percentage_protected", "region_100km_cultivated", "share_of_population_within_400m_of_open_space", "rainfall_monthly_max", "city_ndvi", "temperature_monthly_max")])
create_fifty_rows_of_oob(merlin_city_data_fixed[,c("response", "merlin_pool_size", "biome_name", "realm", "region_100km_ssm", "temperature_annual_average", "temperature_monthly_min", "region_50km_elevation_delta",  "rainfall_monthly_min", "permanent_water", "region_50km_susm", "region_20km_ndvi", "region_20km_urban", "shrubs",  "city_gdp_per_population", "happiness_positive_effect", "city_percentage_protected", "region_100km_cultivated", "share_of_population_within_400m_of_open_space", "rainfall_monthly_max", "city_ndvi", "temperature_monthly_max", "city_average_pop_density")])
create_fifty_rows_of_oob(merlin_city_data_fixed[,c("response", "merlin_pool_size", "biome_name", "realm", "region_100km_ssm", "temperature_annual_average", "temperature_monthly_min", "region_50km_elevation_delta",  "rainfall_monthly_min", "permanent_water", "region_50km_susm", "region_20km_ndvi", "region_20km_urban", "shrubs",  "city_gdp_per_population", "happiness_positive_effect", "city_percentage_protected", "region_100km_cultivated", "share_of_population_within_400m_of_open_space", "rainfall_monthly_max", "city_ndvi", "temperature_monthly_max", "city_average_pop_density", "region_50km_average_pop_density")])
create_fifty_rows_of_oob(merlin_city_data_fixed[,c("response", "merlin_pool_size", "biome_name", "realm", "region_100km_ssm", "temperature_annual_average", "temperature_monthly_min", "region_50km_elevation_delta",  "rainfall_monthly_min", "permanent_water", "region_50km_susm", "region_20km_ndvi", "region_20km_urban", "shrubs",  "city_gdp_per_population", "happiness_positive_effect", "city_percentage_protected", "region_100km_cultivated", "share_of_population_within_400m_of_open_space", "rainfall_monthly_max", "city_ndvi", "temperature_monthly_max", "city_average_pop_density", "region_50km_average_pop_density", "rainfall_annual_average")])
```

"merlin_pool_size", "biome_name", "realm"


```{r}
birdlife_city_data <- fetch_city_data_for('birdlife')
birdlife_city_data
```

```{r}
ggplot(birdlife_city_data, aes(response)) + geom_histogram(binwidth = 1)
```

```{r}
birdlife_city_data_fixed <- rfImpute(response ~ ., birdlife_city_data)
birdlife_city_data_fixed
```

```{r}
select_variables_from_random_forest(birdlife_city_data_fixed)
```

```{r}
exclude_birdlife <- !names(birdlife_city_data_fixed) %in% select_scales(urban = 100, cultivated = 100, elevation_delta = 20, mean_elevation = 100, average_pop_density = 20, includes_estuary = NA, ssm = 50, susm = 100, ndvi = 100, percentage_protected = 100)

birdlife_city_data_fixed_single_scale <- birdlife_city_data_fixed[,exclude_birdlife]
birdlife_city_data_fixed_single_scale
```

```{r}
select_variables_from_random_forest(birdlife_city_data_fixed_single_scale)
```

```{r}
create_fifty_rows_of_oob(birdlife_city_data_fixed[,c("response", "population_growth")])
create_fifty_rows_of_oob(birdlife_city_data_fixed[,c("response", "population_growth", "region_50km_ssm")])
create_fifty_rows_of_oob(birdlife_city_data_fixed[,c("response", "population_growth", "region_50km_ssm", "birdlife_pool_size")])
create_fifty_rows_of_oob(birdlife_city_data_fixed[,c("response", "population_growth", "region_50km_ssm", "birdlife_pool_size", "biome_name")])
create_fifty_rows_of_oob(birdlife_city_data_fixed[,c("response", "population_growth", "region_50km_ssm", "birdlife_pool_size", "biome_name", "region_100km_cultivated")])
create_fifty_rows_of_oob(birdlife_city_data_fixed[,c("response", "population_growth", "region_50km_ssm", "birdlife_pool_size", "biome_name", "region_100km_cultivated", "city_ndvi")])
create_fifty_rows_of_oob(birdlife_city_data_fixed[,c("response", "population_growth", "region_50km_ssm", "birdlife_pool_size", "biome_name", "region_100km_cultivated", "city_ndvi", "temperature_monthly_min")])
create_fifty_rows_of_oob(birdlife_city_data_fixed[,c("response", "population_growth", "region_50km_ssm", "birdlife_pool_size", "biome_name", "region_100km_cultivated", "city_ndvi", "temperature_monthly_min", "percentage_urban_area_as_open_public_spaces")])
create_fifty_rows_of_oob(birdlife_city_data_fixed[,c("response", "population_growth", "region_50km_ssm", "birdlife_pool_size", "biome_name", "region_100km_cultivated", "city_ndvi", "temperature_monthly_min", "percentage_urban_area_as_open_public_spaces", "rainfall_monthly_min")])
create_fifty_rows_of_oob(birdlife_city_data_fixed[,c("response", "population_growth", "region_50km_ssm", "birdlife_pool_size", "biome_name", "region_100km_cultivated", "city_ndvi", "temperature_monthly_min", "percentage_urban_area_as_open_public_spaces", "rainfall_monthly_min", "region_100km_susm")])
create_fifty_rows_of_oob(birdlife_city_data_fixed[,c("response", "population_growth", "region_50km_ssm", "birdlife_pool_size", "biome_name", "region_100km_cultivated", "city_ndvi", "temperature_monthly_min", "percentage_urban_area_as_open_public_spaces", "rainfall_monthly_min", "region_100km_susm", "region_20km_average_pop_density")])
create_fifty_rows_of_oob(birdlife_city_data_fixed[,c("response", "population_growth", "region_50km_ssm", "birdlife_pool_size", "biome_name", "region_100km_cultivated", "city_ndvi", "temperature_monthly_min", "percentage_urban_area_as_open_public_spaces", "rainfall_monthly_min", "region_100km_susm", "region_20km_average_pop_density", "city_ssm")])
create_fifty_rows_of_oob(birdlife_city_data_fixed[,c("response", "population_growth", "region_50km_ssm", "birdlife_pool_size", "biome_name", "region_100km_cultivated", "city_ndvi", "temperature_monthly_min", "percentage_urban_area_as_open_public_spaces", "rainfall_monthly_min", "region_100km_susm", "region_20km_average_pop_density", "city_ssm", "permanent_water")])
create_fifty_rows_of_oob(birdlife_city_data_fixed[,c("response", "population_growth", "region_50km_ssm", "birdlife_pool_size", "biome_name", "region_100km_cultivated", "city_ndvi", "temperature_monthly_min", "percentage_urban_area_as_open_public_spaces", "rainfall_monthly_min", "region_100km_susm", "region_20km_average_pop_density", "city_ssm", "permanent_water", "rainfall_monthly_max")])
create_fifty_rows_of_oob(birdlife_city_data_fixed[,c("response", "population_growth", "region_50km_ssm", "birdlife_pool_size", "biome_name", "region_100km_cultivated", "city_ndvi", "temperature_monthly_min", "percentage_urban_area_as_open_public_spaces", "rainfall_monthly_min", "region_100km_susm", "region_20km_average_pop_density", "city_ssm", "permanent_water", "rainfall_monthly_max", "region_100km_urban")])
create_fifty_rows_of_oob(birdlife_city_data_fixed[,c("response", "population_growth", "region_50km_ssm", "birdlife_pool_size", "biome_name", "region_100km_cultivated", "city_ndvi", "temperature_monthly_min", "percentage_urban_area_as_open_public_spaces", "rainfall_monthly_min", "region_100km_susm", "region_20km_average_pop_density", "city_ssm", "permanent_water", "rainfall_monthly_max", "region_100km_urban", "temperature_annual_average")])
create_fifty_rows_of_oob(birdlife_city_data_fixed[,c("response", "population_growth", "region_50km_ssm", "birdlife_pool_size", "biome_name", "region_100km_cultivated", "city_ndvi", "temperature_monthly_min", "percentage_urban_area_as_open_public_spaces", "rainfall_monthly_min", "region_100km_susm", "region_20km_average_pop_density", "city_ssm", "permanent_water", "rainfall_monthly_max", "region_100km_urban", "temperature_annual_average", "percentage_urban_area_as_open_public_spaces_and_streets")])
create_fifty_rows_of_oob(birdlife_city_data_fixed[,c("response", "population_growth", "region_50km_ssm", "birdlife_pool_size", "biome_name", "region_100km_cultivated", "city_ndvi", "temperature_monthly_min", "percentage_urban_area_as_open_public_spaces", "rainfall_monthly_min", "region_100km_susm", "region_20km_average_pop_density", "city_ssm", "permanent_water", "rainfall_monthly_max", "region_100km_urban", "temperature_annual_average", "percentage_urban_area_as_open_public_spaces_and_streets", "region_20km_elevation_delta")])
create_fifty_rows_of_oob(birdlife_city_data_fixed[,c("response", "population_growth", "region_50km_ssm", "birdlife_pool_size", "biome_name", "region_100km_cultivated", "city_ndvi", "temperature_monthly_min", "percentage_urban_area_as_open_public_spaces", "rainfall_monthly_min", "region_100km_susm", "region_20km_average_pop_density", "city_ssm", "permanent_water", "rainfall_monthly_max", "region_100km_urban", "temperature_annual_average", "percentage_urban_area_as_open_public_spaces_and_streets", "region_20km_elevation_delta", "share_of_population_within_400m_of_open_space")])
create_fifty_rows_of_oob(birdlife_city_data_fixed[,c("response", "population_growth", "region_50km_ssm", "birdlife_pool_size", "biome_name", "region_100km_cultivated", "city_ndvi", "temperature_monthly_min", "percentage_urban_area_as_open_public_spaces", "rainfall_monthly_min", "region_100km_susm", "region_20km_average_pop_density", "city_ssm", "permanent_water", "rainfall_monthly_max", "region_100km_urban", "temperature_annual_average", "percentage_urban_area_as_open_public_spaces_and_streets", "region_20km_elevation_delta", "share_of_population_within_400m_of_open_space", "shrubs")])
create_fifty_rows_of_oob(birdlife_city_data_fixed[,c("response", "population_growth", "region_50km_ssm", "birdlife_pool_size", "biome_name", "region_100km_cultivated", "city_ndvi", "temperature_monthly_min", "percentage_urban_area_as_open_public_spaces", "rainfall_monthly_min", "region_100km_susm", "region_20km_average_pop_density", "city_ssm", "permanent_water", "rainfall_monthly_max", "region_100km_urban", "temperature_annual_average", "percentage_urban_area_as_open_public_spaces_and_streets", "region_20km_elevation_delta", "share_of_population_within_400m_of_open_space", "shrubs", "mean_population_exposure_to_pm2_5_2019")])
create_fifty_rows_of_oob(birdlife_city_data_fixed[,c("response", "population_growth", "region_50km_ssm", "birdlife_pool_size", "biome_name", "region_100km_cultivated", "city_ndvi", "temperature_monthly_min", "percentage_urban_area_as_open_public_spaces", "rainfall_monthly_min", "region_100km_susm", "region_20km_average_pop_density", "city_ssm", "permanent_water", "rainfall_monthly_max", "region_100km_urban", "temperature_annual_average", "percentage_urban_area_as_open_public_spaces_and_streets", "region_20km_elevation_delta", "share_of_population_within_400m_of_open_space", "shrubs", "mean_population_exposure_to_pm2_5_2019", "city_average_pop_density")])
create_fifty_rows_of_oob(birdlife_city_data_fixed[,c("response", "population_growth", "region_50km_ssm", "birdlife_pool_size", "biome_name", "region_100km_cultivated", "city_ndvi", "temperature_monthly_min", "percentage_urban_area_as_open_public_spaces", "rainfall_monthly_min", "region_100km_susm", "region_20km_average_pop_density", "city_ssm", "permanent_water", "rainfall_monthly_max", "region_100km_urban", "temperature_annual_average", "percentage_urban_area_as_open_public_spaces_and_streets", "region_20km_elevation_delta", "share_of_population_within_400m_of_open_space", "shrubs", "mean_population_exposure_to_pm2_5_2019", "city_average_pop_density", "percentage_urban_area_as_streets")])
create_fifty_rows_of_oob(birdlife_city_data_fixed[,c("response", "population_growth", "region_50km_ssm", "birdlife_pool_size", "biome_name", "region_100km_cultivated", "city_ndvi", "temperature_monthly_min", "percentage_urban_area_as_open_public_spaces", "rainfall_monthly_min", "region_100km_susm", "region_20km_average_pop_density", "city_ssm", "permanent_water", "rainfall_monthly_max", "region_100km_urban", "temperature_annual_average", "percentage_urban_area_as_open_public_spaces_and_streets", "region_20km_elevation_delta", "share_of_population_within_400m_of_open_space", "shrubs", "mean_population_exposure_to_pm2_5_2019", "city_average_pop_density", "percentage_urban_area_as_streets", "rainfall_annual_average")])
create_fifty_rows_of_oob(birdlife_city_data_fixed[,c("response", "population_growth", "region_50km_ssm", "birdlife_pool_size", "biome_name", "region_100km_cultivated", "city_ndvi", "temperature_monthly_min", "percentage_urban_area_as_open_public_spaces", "rainfall_monthly_min", "region_100km_susm", "region_20km_average_pop_density", "city_ssm", "permanent_water", "rainfall_monthly_max", "region_100km_urban", "temperature_annual_average", "percentage_urban_area_as_open_public_spaces_and_streets", "region_20km_elevation_delta", "share_of_population_within_400m_of_open_space", "shrubs", "mean_population_exposure_to_pm2_5_2019", "city_average_pop_density", "percentage_urban_area_as_streets", "rainfall_annual_average", "city_susm")])
create_fifty_rows_of_oob(birdlife_city_data_fixed[,c("response", "population_growth", "region_50km_ssm", "birdlife_pool_size", "biome_name", "region_100km_cultivated", "city_ndvi", "temperature_monthly_min", "percentage_urban_area_as_open_public_spaces", "rainfall_monthly_min", "region_100km_susm", "region_20km_average_pop_density", "city_ssm", "permanent_water", "rainfall_monthly_max", "region_100km_urban", "temperature_annual_average", "percentage_urban_area_as_open_public_spaces_and_streets", "region_20km_elevation_delta", "share_of_population_within_400m_of_open_space", "shrubs", "mean_population_exposure_to_pm2_5_2019", "city_average_pop_density", "percentage_urban_area_as_streets", "rainfall_annual_average", "city_susm", "region_100km_ndvi")])
create_fifty_rows_of_oob(birdlife_city_data_fixed[,c("response", "population_growth", "region_50km_ssm", "birdlife_pool_size", "biome_name", "region_100km_cultivated", "city_ndvi", "temperature_monthly_min", "percentage_urban_area_as_open_public_spaces", "rainfall_monthly_min", "region_100km_susm", "region_20km_average_pop_density", "city_ssm", "permanent_water", "rainfall_monthly_max", "region_100km_urban", "temperature_annual_average", "percentage_urban_area_as_open_public_spaces_and_streets", "region_20km_elevation_delta", "share_of_population_within_400m_of_open_space", "shrubs", "mean_population_exposure_to_pm2_5_2019", "city_average_pop_density", "percentage_urban_area_as_streets", "rainfall_annual_average", "city_susm", "region_100km_ndvi", "happiness_future_life")])

```

"population_growth", "region_50km_ssm", "birdlife_pool_size"



------------------------------------------
So....
------------------------------------------
Merlin: "merlin_pool_size", "biome_name", "realm"
Birdlife: "population_growth", "region_50km_ssm", "birdlife_pool_size"



-----------------------------
Try Modelling
-----------------------------

```{r}
library(boot)
```



```{r}
merlin_city_data_named <- fetch_city_data_for('merlin', T)
birdlife_city_data_named <- fetch_city_data_for('birdlife', T)
```

------------------------------------------------------------------
Use cross validation and dropping terms to find best model
------------------------------------------------------------------

```{r}
cv.glm(merlin_city_data_fixed_no_boreal, glm(formula = response ~ merlin_pool_size + population_growth + region_50km_ssm + biome_name + realm, data = merlin_city_data_fixed_no_boreal))$delta[1]
```

-- Can we drop one?

```{r}
cv.glm(merlin_city_data_fixed_no_boreal, glm(formula = response ~ merlin_pool_size + population_growth + region_50km_ssm + biome_name, data = merlin_city_data_fixed_no_boreal))$delta[1]
```

```{r}
cv.glm(merlin_city_data_fixed_no_boreal, glm(formula = response ~ merlin_pool_size + population_growth + region_50km_ssm + realm, data = merlin_city_data_fixed_no_boreal))$delta[1]
```

```{r}
cv.glm(merlin_city_data_fixed_no_boreal, glm(formula = response ~ merlin_pool_size + population_growth + biome_name + realm, data = merlin_city_data_fixed_no_boreal))$delta[1]
```

```{r}
cv.glm(merlin_city_data_fixed_no_boreal, glm(formula = response ~ merlin_pool_size + region_50km_ssm + biome_name + realm, data = merlin_city_data_fixed_no_boreal))$delta[1]
```

```{r}
cv.glm(merlin_city_data_fixed_no_boreal, glm(formula = response ~ region_50km_ssm + biome_name + realm, data = merlin_city_data_fixed_no_boreal))$delta[1]
```

-- drop populaion_growth (14.30384)

-- can we drop another?
```{r}
cv.glm(merlin_city_data_fixed_no_boreal, glm(formula = response ~ merlin_pool_size + region_50km_ssm + biome_name, data = merlin_city_data_fixed_no_boreal))$delta[1]
```

```{r}
cv.glm(merlin_city_data_fixed_no_boreal, glm(formula = response ~ merlin_pool_size + region_50km_ssm + realm, data = merlin_city_data_fixed_no_boreal))$delta[1]
```

```{r}
cv.glm(merlin_city_data_fixed_no_boreal, glm(formula = response ~ merlin_pool_size + biome_name + realm, data = merlin_city_data_fixed_no_boreal))$delta[1]
```

```{r}
cv.glm(merlin_city_data_fixed_no_boreal, glm(formula = response ~ region_50km_ssm + biome_name + realm, data = merlin_city_data_fixed_no_boreal))$delta[1]
```

-- drop ssm (14.20889)

-- can we drop another?

```{r}
cv.glm(merlin_city_data_fixed_no_boreal, glm(formula = response ~ merlin_pool_size + biome_name, data = merlin_city_data_fixed_no_boreal))$delta[1]
```

```{r}
cv.glm(merlin_city_data_fixed_no_boreal, glm(formula = response ~ merlin_pool_size + realm, data = merlin_city_data_fixed_no_boreal))$delta[1]
```

```{r}
cv.glm(merlin_city_data_fixed_no_boreal, glm(formula = response ~ biome_name + realm, data = merlin_city_data_fixed_no_boreal))$delta[1]
```

-- drop biome (13.10131)

```{r}
cv.glm(merlin_city_data_fixed_no_boreal, glm(formula = response ~ merlin_pool_size, data = merlin_city_data_fixed_no_boreal))$delta[1]
```

```{r}
cv.glm(merlin_city_data_fixed_no_boreal, glm(formula = response ~ realm, data = merlin_city_data_fixed_no_boreal))$delta[1]
```

```{r}
cv.glm(merlin_city_data_fixed_no_boreal, glm(formula = response ~ merlin_pool_size * realm, data = merlin_city_data_fixed_no_boreal))$delta[1]
```

-----------------------------------------------
-- best model with merlin is pool size + realm (CV error 13.10131)
-----------------------------------------------
```{r}
summary(glm(data = merlin_city_data_fixed, formula = response ~ merlin_pool_size + realm))
```


```{r}
cv.glm(birdlife_city_data_fixed_no_boreal, glm(formula = response ~ birdlife_pool_size + population_growth + region_50km_ssm + biome_name + realm, data = birdlife_city_data_fixed_no_boreal))$delta[1]
```

-- can we drop a variable?

```{r}
cv.glm(birdlife_city_data_fixed_no_boreal, glm(formula = response ~ birdlife_pool_size + population_growth + region_50km_ssm + biome_name, data = birdlife_city_data_fixed_no_boreal))$delta[1]
```

```{r}
cv.glm(birdlife_city_data_fixed_no_boreal, glm(formula = response ~ birdlife_pool_size + population_growth + region_50km_ssm + realm, data = birdlife_city_data_fixed_no_boreal))$delta[1]
```

```{r}
cv.glm(birdlife_city_data_fixed_no_boreal, glm(formula = response ~ birdlife_pool_size + population_growth + biome_name + realm, data = birdlife_city_data_fixed_no_boreal))$delta[1]
```

```{r}
cv.glm(birdlife_city_data_fixed_no_boreal, glm(formula = response ~ birdlife_pool_size + region_50km_ssm + biome_name + realm, data = birdlife_city_data_fixed_no_boreal))$delta[1]
```

```{r}
cv.glm(birdlife_city_data_fixed_no_boreal, glm(formula = response ~ population_growth + region_50km_ssm + biome_name + realm, data = birdlife_city_data_fixed_no_boreal))$delta[1]
```

-- drop biome (5.492536)

```{r}
cv.glm(birdlife_city_data_fixed_no_boreal, glm(formula = response ~ birdlife_pool_size + population_growth + region_50km_ssm, data = birdlife_city_data_fixed_no_boreal))$delta[1]
```

```{r}
cv.glm(birdlife_city_data_fixed_no_boreal, glm(formula = response ~ birdlife_pool_size + population_growth + realm, data = birdlife_city_data_fixed_no_boreal))$delta[1]
```

```{r}
cv.glm(birdlife_city_data_fixed_no_boreal, glm(formula = response ~ birdlife_pool_size + region_50km_ssm + realm, data = birdlife_city_data_fixed_no_boreal))$delta[1]
```

```{r}
cv.glm(birdlife_city_data_fixed_no_boreal, glm(formula = response ~ population_growth + region_50km_ssm + realm, data = birdlife_city_data_fixed_no_boreal))$delta[1]
```

-- drop population growth (5.38609)
```{r}
cv.glm(birdlife_city_data_fixed_no_boreal, glm(formula = response ~ birdlife_pool_size + region_50km_ssm, data = birdlife_city_data_fixed_no_boreal))$delta[1]
```

```{r}
cv.glm(birdlife_city_data_fixed_no_boreal, glm(formula = response ~ birdlife_pool_size + realm, data = birdlife_city_data_fixed_no_boreal))$delta[1]
```

```{r}
cv.glm(birdlife_city_data_fixed_no_boreal, glm(formula = response ~ region_50km_ssm + realm, data = birdlife_city_data_fixed_no_boreal))$delta[1]
```

-- drop ssm (5.38033)

```{r}
cv.glm(birdlife_city_data_fixed_no_boreal, glm(formula = response ~ birdlife_pool_size, data = birdlife_city_data_fixed_no_boreal))$delta[1]
```

```{r}
cv.glm(birdlife_city_data_fixed_no_boreal, glm(formula = response ~ realm, data = birdlife_city_data_fixed_no_boreal))$delta[1]
```

```{r}
cv.glm(birdlife_city_data_fixed_no_boreal, glm(formula = response ~ birdlife_pool_size * realm, data = birdlife_city_data_fixed_no_boreal))$delta[1]
```
----------------------------------------------------
-- so best model with birdlife is pool size + realm
----------------------------------------------------

```{r}
summary(glm(data = birdlife_city_data_fixed, formula = response ~ birdlife_pool_size + realm))
```

```{r}
ggplot(merlin_city_data_fixed_no_boreal, aes(x = merlin_pool_size, y = response)) + geom_point() + geom_smooth(method = "glm", se = F) + facet_wrap(~ realm)
```

```{r}
ggplot(birdlife_city_data_fixed_no_boreal, aes(x = birdlife_pool_size, y = response)) + geom_point() + geom_smooth(method = "glm", se = F) + facet_wrap(~ realm)
```

```{r}
birdlife.fit <- glm(data = birdlife_city_data_fixed_no_boreal, formula = response ~ birdlife_pool_size + realm)
summary(birdlife.fit)
with(summary(birdlife.fit), 1 - deviance/null.deviance)
plot(birdlife.fit)
```

```{r}
birdlife_city_data_fixed_no_boreal[c(16, 53, 116, 123), c("name", "birdlife_pool_size")]
```

```{r}
ggplot(birdlife_city_data_fixed_no_boreal, aes(x = birdlife_pool_size, y = response)) + 
  geom_point(size=1) + 
  geom_smooth(method = "glm", se = F) +
  geom_text(aes(label = name), data = birdlife_city_data_fixed_no_boreal[c(16, 53, 123),], size = 3, position = "dodge", vjust = "inward", hjust = "inward", color = "red", angle=-15) +
  geom_point(data = birdlife_city_data_fixed_no_boreal[c(16, 53, 123),], color = "red") +
  facet_wrap(~ realm) +
  theme_bw() +
  ylab("City Random Effect Intercept") + xlab("Regional Pool Size") + labs(title = "Birdlife")

ggsave("city_effect_richness_birdlife.jpg")
```

```{r}
merlin.fit <- glm(data = merlin_city_data_fixed_no_boreal, formula = response ~ merlin_pool_size + realm)
summary(merlin.fit)
with(summary(merlin.fit), 1 - deviance/null.deviance)
plot(merlin.fit)
```

```{r}
city_data[c(7, 24, 77), c("name", "birdlife_pool_size", "merlin_pool_size")]
```

```{r}
ggplot(merlin_city_data_fixed_no_boreal, aes(x = merlin_pool_size, y = response)) + 
  geom_point(size = 1) + 
  geom_smooth(method = "glm", se = F) +
  geom_text(aes(label = name), data = merlin_city_data_fixed_no_boreal[c(7, 24, 77),], size = 3, position = "dodge", vjust = "inward", hjust = "inward", color = "red", angle=-15) +
  geom_point(data = merlin_city_data_fixed_no_boreal[c(7, 24, 77),], color = "red") +
  facet_wrap(~ realm) +
  theme_bw() +
  ylab("City Random Effect Intercept") + xlab("Regional Pool Size") + labs(title = "eBird")

ggsave("city_effect_richness_merlin.jpg")
```

```{r}
merlin_city_data_fixed_no_boreal$residuals <- resid(merlin.fit)
ggplot(merlin_city_data_fixed_no_boreal, aes(y = response, x = residuals)) + 
  geom_smooth(method = "lm", se = F) +
  geom_point(aes(color = realm)) + 
  geom_text(aes(label = name), data = merlin_city_data_fixed_no_boreal[c(7, 24, 77),], size = 4, position = "dodge", vjust = "inward", hjust = "inward") + 
  labs(title = "Merlin", subtitle = paste("Correlation", cor(merlin_city_data_fixed_no_boreal$residuals, merlin_city_data_fixed_no_boreal$response))) +
  theme_bw()
```

```{r}
birdlife_city_data_fixed_no_boreal$residuals <- resid(birdlife.fit)
ggplot(birdlife_city_data_fixed_no_boreal, aes(y = response, x = residuals)) + 
  geom_smooth(method = "lm", se = F, alpha = 0.5) +
  geom_point(aes(color = realm)) + 
  geom_text(aes(label = name), data = birdlife_city_data_fixed_no_boreal[c(16, 53, 124),], size = 4, position = "dodge", vjust = "inward", hjust = "inward") +
  labs(title = "Birdlife", subtitle = paste("Correlation", cor(birdlife_city_data_fixed_no_boreal$residuals, birdlife_city_data_fixed_no_boreal$response))) +
  theme_bw()
```

```{r}
birdlife_city_data_fixed_no_boreal$residuals_of_fit <- resid(lm(response ~ residuals, birdlife_city_data_fixed_no_boreal))
ggplot(birdlife_city_data_fixed_no_boreal, aes(x = realm, y = residuals_of_fit)) + geom_boxplot()
```

```{r}
birdlife.aov <- aov(birdlife_city_data_fixed_no_boreal$residuals_of_fit ~ birdlife_city_data_fixed_no_boreal$realm)
summary(birdlife.aov)
```

```{r}
merlin_city_data_fixed_no_boreal$residuals_of_fit <- resid(lm(response ~ residuals, merlin_city_data_fixed_no_boreal))
ggplot(merlin_city_data_fixed_no_boreal, aes(x = realm, y = residuals_of_fit)) + geom_boxplot()
```

```{r}
merlin.aov <- aov(merlin_city_data_fixed_no_boreal$residuals_of_fit ~ merlin_city_data_fixed_no_boreal$realm)
summary(merlin.aov)
```


```{r}
ggplot(merlin_city_data_fixed_no_boreal, aes(residuals)) + geom_histogram(binwidth = 1)
```

```{r}
ggplot(birdlife_city_data_fixed_no_boreal, aes(residuals)) + geom_histogram(binwidth = 1)
```

```{r}
shapiro.test(birdlife_city_data_fixed_no_boreal$response)
shapiro.test(birdlife_city_data_fixed_no_boreal$residuals)
shapiro.test(birdlife_city_data_fixed_no_boreal$residuals_of_fit)

shapiro.test(merlin_city_data_fixed_no_boreal$response)
shapiro.test(merlin_city_data_fixed_no_boreal$residuals)
shapiro.test(merlin_city_data_fixed_no_boreal$residuals_of_fit)
```

Variances of groups are NOT significantly different:
```{r}
bartlett.test(merlin_city_data_fixed_no_boreal$residuals, merlin_city_data_fixed_no_boreal$realm) 
leveneTest(merlin_city_data_fixed_no_boreal$residuals, merlin_city_data_fixed_no_boreal$realm) 
```

Variances of groups are NOT significantly different:
```{r}
bartlett.test(birdlife_city_data_fixed_no_boreal$residuals, birdlife_city_data_fixed_no_boreal$realm) 
leveneTest(birdlife_city_data_fixed_no_boreal$residuals, birdlife_city_data_fixed_no_boreal$realm) 
```

Variances of groups are significantly different:
```{r}
bartlett.test(merlin_city_data_fixed_no_boreal$residuals_of_fit, merlin_city_data_fixed_no_boreal$realm) 
leveneTest(merlin_city_data_fixed_no_boreal$residuals_of_fit, merlin_city_data_fixed_no_boreal$realm) 
```

Variances of groups are significantly different:
```{r}
bartlett.test(birdlife_city_data_fixed_no_boreal$residuals_of_fit, birdlife_city_data_fixed_no_boreal$realm) 
leveneTest(birdlife_city_data_fixed_no_boreal$residuals_of_fit, birdlife_city_data_fixed_no_boreal$realm) 
```


-------------------------
Check AIC
-------------------------

```{r}
AIC(
  glm(data = merlin_city_data_fixed, formula = response ~ merlin_pool_size + population_growth + region_50km_ssm + biome_name + realm),
  glm(data = merlin_city_data_fixed, formula = response ~ merlin_pool_size + realm)
)
```

```{r}
AIC(
  glm(data = birdlife_city_data_fixed, formula = response ~ birdlife_pool_size + population_growth + region_50km_ssm + biome_name + realm),
  glm(data = birdlife_city_data_fixed, formula = response ~ birdlife_pool_size + realm)
)
```
